Cycles of the logistic map

### 2013-06-06 doc/ijbc/response.txt

Dear Editor,

Thank you and the referees for extremely helpful suggestions and 
comments.  I have carefully modified the manuscript according to your 
suggestions and the referee reports. Below are the responses to the 
points raised by the referees. Additionally, the cycle polynomials 
for the logistic map have been computed for n = 14.

Regards,
Cheng

Response to Reviewer 1

Reviewer #1: There are several methods to compute these polynomials 
already in the literature: Kotsireas and Karamanos's Groebner base 
approach, Burm and Fishback's resultant approach, and Blackhurst's 
rational canonical matrix approach. (Bailey and Broadhurst's 
integer-relation detection approach could also be included, but their 
motivation was more to push the limits of their algorithm and less to 
shed light on chaos theory.) Adding one more method to these other 
methods doesn't hurt, but it's not really clear from the paper what 
this new approach contributes beyond Burm and Fishback's or 
Blackhurst's methods. Section 5 has a comparison between this new 
method and Kotsireas and Karamanos's approach, which is good, but the 
comparison could be improved. Instead of listing run times---which can 
be affected by such things as the hardware specifications, the 
algorithm being used (e.g, Buchburger's versus Faguere's F4 alorithm), 
and implementation details---it would be better to
give computational complexities. Since Groebner base algorithms are 
doubly exponential in the number of variables, it's not surprising this 
new method performs better. But how does it compare to Burm and 
Fishback's and Blackhurst's methods? What are the computational 
complexities of those methods and this new method? One advantage of 
Burm and Fishback's method, complexity issues aside, is the ease of 
programming: most computer algebra system have built-in commands to 
iterate functions and take resultants, but not to calculate cyclic 
polynomials. Does this new method also have reasons, beyond complexity 
issues, to prefer it over other methods?  In summary, section 5 needs 
to be expanded to more fully compare and contrast this new method with 
previous methods. At the very least, the computational complexities for 
this new method, Kotsireas and Karamanos's approach, Burm and 
Fishback's approach, and Blackurst's approach should be given.

Response:  First, I have implemented both the Grobner basis approach by 
Kotsireas and Karamanos and the resultant approach by Burm and Fishback 
using both Magma and Mathematica.  The results of two competing methods 
are compared with the current method in the new Fig. 1 of the current 
manuscript.  Clearly, for n >= 8, the present method are two orders of 
magnitudes more efficient than the competitors.  Further, Magma is 
known to be more optimized than Mathematica, and it is especially 
famous for its fast Grobner basis routine.  This is supported by Fig. 
1, the Grobner bases and resultants took much less time to compute 
using Magma.  Even so, our method, only implemented in a high-level 
language Mathematica, is still much faster.  This demonstrate the 
practical efficiency of the method.  Below I will further discuss that 
the underlying reason for the performance gain is linked to the use of 
the cyclic structure of the equations.

Secondly,  the current method is expected to outperform the resultant 
method.  The resultant of two polynomials is defined as the determinant 
of the Sylvester matrix of the two polynomials.  The side of the 
Sylvester matrix, which is square, is equal to the sum of the degrees 
of the two polynomials.  In this case,  the two polynomials are 
	f^n(x) – x,
and
            (f^n)'(x) -/+ 1.
It is easy to see that the matrix roughly has N = 2^(n+1) rows/columns. 
 Now the key step of our method is also the computation of a 
determinant, but it is the one from the characteristic equation.  Now 
our matrix only has roughly N = 2^n/n rows/columns.  So, even without 
specifying the underlying algorithm of computing the determinant, we 
expect the current method to outperform the resultant method.  This 
point is now added in Section 2.9 in the revised manuscript.  If we 
assume the Bareiss algorithm for the determinant, then the run time 
goes as N^3 ~ N^4, or 8^n ~ 16^n.  This estimate agrees with the 
results shown in Fig. 1.  However, the underlying routine of computing 
the determinant used by Mathematica or Magma may differ greatly in 
algorithm and implementation, therefore the specific algorithms are not 
discussed in the manuscript for brevity.  In short, the underlying 
reason for the performance gain is that the current algorithm exploits 
the cyclic structures of the equations, which greatly reduces the 
matrix size.

Thirdly, the Grobner basis method is expected to perform similarly to 
the resultant method.  This is supported by Fig. 1.  Conceptually, the 
resultant step can be considered as the construction of the final basic 
polynomial between x1 and R, after other x2, x3, ..., xn are 
eliminated.  The conceptual relation is briefly discussed in Section 
2.1 in the revised manuscript.  For this problem, the computational 
times of both methods grow exponentially with n, which is fortuitous, 
as it is expected to have worse scaling for a general problem of n 
variables as suggested by the referee.

On the other hand, Blackhurst's method, according to my understanding 
of the cited reference, is not an independent method that computes the 
cycle polynomials.  Instead, it is a formula that computes the *smaller 
factors* of the onset polynomial of the n-cycles.  So it needs another 
method to first compute the onset polynomial of n-cycles itself.  At 
the onset point, the polynomial of n-cycles can be factorized due to 
the intersection between the n-cycles and the shorter d-cycles (as 
discussed in the Section 2.8 of the current version). These 
intersections contribute smaller factors to the n-cycle polynomial.  
For example, in case of n = 4, the onset polynomial is
	P4(onset)  =  (5-4R) [(4R+1)^2 + 4] [108 – (4R-3)^3],
and the smaller factors are 5-4R (from the 2-cycles that are about to 
double their period) and (4R+1)^2 + 4 (from 1-cycles).  In the cited 
reference, Blackhurst showed a formula of finding the smaller factors 
in terms of canonical matrices.  The formula requires one to 
numerically compute the cycle points explicitly.  Alternatively, we 
show here that the factors can be compactly expressed as a cyclotomic 
product in the denominator of Eq. (13) of the current manuscript.  The 
latter approach is easier to use for it does not require the explicit 
computation of the cycle points at different R values (or the \lambda 
values in Blackhurst's paper) .  The relation between the two formulas 
are now pointed out in this version.  It is my understanding that 
Blackhurst's canonical-matrix formula requires *a priori* the d-cycle 
and n-cycle polynomials as well as the cycle points, which can be very 
difficult to obtain for a large n.  There may be, however, some 
unpublished ongoing work that I am not aware of.  If this is so, I am 
happy to add a reference.

Response to Reviewer 2

Reviewer #2: The paper is well written. However, the length could be 
greatly reduced. Most figures and tables are not so informative, so can 
be removed. Some examples can also be deleted. Derivations related to 
cycle counting seem not new, so the author may just give the results 
with some brief description and relevant references.

Response:  I have removed all previous figures and tables, minimized 
the derivations on the cycle counting, and simplified the examples.  
The overall length has been cut over a half, and the article looks more 
compact now.  Thank you for the suggestions.

### 2013-06-06 doc/pc-ijbc.tex
add response.txt

Since the $n$-cycles points constitute
  a subset of the fixed points of $f^n$,
we will first find
  the polynomials at the stability boundaries
  for the fixed points of $f^n$
  (\refsecs{cyclic}-\ref{sec:examples}).
The polynomials contain
  extraneous factors contributed by short $d$-cycles with $d|n$,
  which are also fixed points of $f^n$.
The factors are removed in \refsec{primfac}.

@@

\refthm{sqrfree}
  shows that any cyclic polynomial can be expanded
  as a combination of the square-free ones.
At the onset and bifurcation points,
  the $\NB$ square-free cyclic polynomials $C_p$ are themselves
  linearly connected by \refeq{der}.
This condition yields a characteristic equation,
  which solves our problem.

@@

%\begin{theorem}
%  The polynomial $P_n\Rlam$ of all $n$-cycles
%  is a factor of the characteristic polynomial $A_n\Rlam$,
%  and can be computed as
  \begin{equation}
    P_n\Rlam
    = \prod_{cd = n} B_d^c\Rlam^{\mu(c)},
  \label{eq:primfac}
  \end{equation}
where
  $B_d^c\Rlam \equiv \prod_{k=1}^c A_{d}(R, e^{2k\pi i/c} \lambda^{1/c})$,
  particularly, $B_n^1\rlam = A_n\rlam$,
  and
  $\lambda^{1/c}$ is a complex $c$th root of $\lambda$;
  the outer product runs over all divisors $d$ of $n$.
%\end{theorem}
%
%
%\begin{remark}
%Here, ``minimal'' means that the polynomial contains all zeros.
%It happens to be irreducible for a general $\lambda$.
%At $\lambda = 1$, the polynomial of $R$, however, can be factorized (see later).
%\end{remark}
%
%\begin{remark}
The M\"obius function $\mu(n)$ is $(-1)^k$
  if $n$ is the product of $k$ distinct primes,
  or 0 if $n$ is divisible by a square of a prime;
%Thus,
$\mu(n) = 1, -1, -1, 0, -1, 1, -1, \ldots,$
  from $n = 1$
%The $\mu(n)$ is useful for inversion:
%$g(n) = \sum_{d|n} \mu(n/d) \, h(d)$
%if and only if $h(n) = \sum_{d|n} g(d)$
 \cite{hardy}.
%\end{remark}
%

@@

First, to count the square-free cyclic polynomials,
  we notice a one-to-one mapping
  from the square-free cyclic polynomials
  to the binary necklaces,
%
which are nonequivalent binary $0$-$1$ strings.
%
Two strings are equivalent if they differ only by a circular shift.
%
For example, for $n = 3$, there are $2^3 = 8$ binary strings,
but only four necklaces: 000, 001, 011 and 111,
since 010 and 100 are equivalent to 001,
so are 110 and 101 to 011.
%
%The mapping is constructed as follows.
%
Now given a square-free cyclic polynomial, if its generator contains $x_k$,
  the $k$th character from the right of the necklace is 1,
  otherwise 0,
e.g. the above four necklaces correspond to
  $C_0$, $C_1$, $C_{12}$ and $C_{123}$, respectively.
%For example, if $n = 3$,
% the square-free polynomials for $000$, $100$, $110$ and $111$ are
%  $C_0 = 1$ (generator: 1),
%  $C_1 = x_1 + x_2 + x_3$ (generator: $x_1$),
%  $C_{12} = x_1 x_2 + x_2 x_3 + x_3 x_1$ (generator: $x_1 x_2$)
%  and $C_{123} = x_1 x_2 x_3$ (generator: $x_1 x_2 x_3$), respectively.
Thus,
%
%\begin{theorem}
the number $\NB(n)$ of the square-free cyclic polynomials
  is equal to the number of length-$n$ binary necklaces,
  which is given by \cite{riordan}
\begin{equation}
  %\deg_\lambda A_n\lam =
  \NB(n) = (1/n) \textstyle\sum_{d|n} \varphi(n/d) 2^{\,d},
\label{eq:necklace}
\end{equation}
%
%\label{thm:necklace}
%\end{theorem}
%
where $\varphi(m)$ is Euler's totient function,
  which counts the number of integers from 1 to $m$
  that are coprime to $m$,
  e.g.,
  $\varphi(1) = 1$,
  $\varphi(3) = 2$ for 1 and 2 are coprime to 3,
  $\varphi(6) = 2$ for 1 and 5 \cite{hardy}.
%
By \refeq{xr},
  $\NB(n) = \deg_\lambda A_n\Rlam$.
The first few $\NB(n)$ are
2, 3, 4, 6, 8, 14, 20, 36, 60, 108, \dots, % 188, 352, 632, \ldots,
  from $n = 1$.

@@

%\begin{theorem}
Next, we define the $n$-cycle polynomial as
\begin{equation}
  P_n\lam \equiv \textstyle\prod_l [
    \Lambda(\vx{n}^{(l)}) - \lambda
  ],
\notag
\end{equation}
where $\vx{n}^{(l)}$ denotes the $l$th $n$-cycle.
%
Thus, $\deg_\lambda P_n\lam$
  is equal to the number $\LB$ of the $n$-cycles,
  which is given by
  \cite{hao, hao2, lutzky}
\begin{equation}
  %\deg_\lambda P_n\lam =
  \LB(n) = (1/n)\textstyle\sum_{d | n} \mu(n/d) 2^{\,d}.
\label{eq:lyndon}
\end{equation}
%\label{thm:lyndon}
%\end{theorem}
%
%
%\end{proof}
%
The first few $\LB(n)$ are
2, 1, 2, 3, 6, 9, 18, 30, 56, 99, \dots %186, 335, 630, \ldots,
from $n = 1$.
%
%
One can verify that
%Incidentally, $\LB(n)$ is also the number of aperiodic necklaces of length $n$
%(e.g., for $n = 4$,
%  ``$0011$'' is aperiodic, but ``$0101$'' is not, owing to the shorter period 2).
%%
%An $n$-necklace is either aperiodic itself or
%  a repeat of an aperiodic $d$-necklace with $d|n$; so
\begin{equation}
  \NB(n) = \textstyle\sum_{d|n} \LB(d).
\notag
%\label{eq:necklacelyndon}
\end{equation}
%
%Since the left side is the degree in $\lambda$ of $A_n\lam$,
Thus, in order for $A_n\lam$ to accommodate all fixed points of $f^n$,
  each distinct $d$-cycle with $d|n$
  contributes one and only one zero to $A_n\lam$.
%
%
Comparing the leading coefficient of $A_n\lam$ leads to
%is unity by \refeq{xr}, we must have
%\begin{theorem}
  \begin{equation}
    A_n\lam = \textstyle\prod_{d|n} Q_d^{n/d}\lam,
    \label{eq:prod}
  \end{equation}
where $Q_d^{n/d}(\lambda) \equiv \prod_l
  \big[\Lambda\big(\vx{n}^{(l)}\big) - \lambda\big]$
  with $\vx{n}^{(l)}$ being the $n$ consecutive points
  in the $l$th $d$-cycle,
%
  or the $d$-cycle points $\vx{d}^{(l)}$
  repeated for $c = n/d$ times:
\begin{equation}
  \vx{n}^{(l)}
  =
  \vx{d}^{(l)} \times c
  =
  \{
    x_1^{(l)}, \ldots, x_d^{(l)},
    \; \ldots, \;
    x_1^{(l)}, \ldots, x_d^{(l)}
  \}.
\notag
\end{equation}
%
It follows that
$\Lambda(\vx{n}) = \left[ \Lambda(\vx{d}) \right]^c$.
%
By factoring $\Lambda^c - \lambda$
  in terms of the $c$ complex roots of $\lambda$,
  we get
%
\begin{align}
Q_d^{c}\lam
%  &= \prod_l
%    \Big\{
%      \lambda - \big[\Lambda\big(\vx{d}^{(l)}\big) \big]^{c}
%    \Big\}
%    \notag \\
  &=
    \prod_l
    \left\{
      \prod_{k=1}^c \,
      \left[
          \Lambda\big(\vx{d}^{(l)}\big)
          -
          e^{2k\pi i/c} \lambda^{1/c}
      \right]
    \right\}
    \notag \\
  &=  \prod_{k=1}^c P_d\big(e^{2k\pi i/c} \lambda^{1/c}\big).
\label{eq:QP}
\end{align}
%
%where we have exchanged the order of the two products in the second step.
%
The radical $\lambda^{1/c}$
  does not prevent
  $Q_d^c\lam$ from being a polynomial of $\lambda$,
  thanks to the invariance
  $\lambda \rightarrow e^{2\pi i} \lambda$;
and
  $\deg_\lambda Q_d^c\lam
    = \deg_\lambda P_d\lam = \LB(d)$.

@@

The most time-consuming step of the algorithm in \refsec{algo}
is the evaluation of the determinant $A_n\Rlam$ in the characteristic
equation (\ref{eq:xr}).
%
When implemented on a computer,
it is helpful to use polynomial interpolation:
we evaluate $A_n\Rlam$ at $\deg_R A_n + 1$ values of $R$,
e.g., $R = 0, \pm1, \pm2,\ldots$, then piece them together
to a polynomial.
%
By using \refeq{primfac}, we only need $\deg_R P_n + 1$ values
  to directly compute $P_n\Rlam$.
%
%
Here,
$\deg_R A_n = \sum_{d|n} \varphi(n/d) 2^{d-1}$,
and $\deg_R P_n = \sum_{d|n} \mu(n/d) 2^{d-1}$
  \cite{mira, stephenson1}.
%
The strategy also allows a trivial parallelization.

We coded the algorithm into a Mathematica program,
and computed the polynomials for $n$ up to 14.
%
As shown in \reffig{logtime}, the algorithm
  outperformed the resultant \cite{burm} and
  Gr\"obner basis \cite{kk1} methods,
  by at least two orders of magnitudes.
%
This is unsurprising,
  because the resultant of $f^n(x)-x$ and $\frac{d}{dx}f^n(x)-\lambda$
  is the determinant of the Sylvester matrix, which has
  $2^{n+1}-1$ rows and columns
  [as $\deg_x f^n(x) = 2^n$],
  while the matrix in \refeq{xr}
  is of only $\NB \approx 2^n/n$ [cf. \refeq{necklace}] rows and columns.
%Thus, the current algorithm is expected to be faster.

@@

Finally, we can further reduce the computational cost
  by half for an even $n$.
%
If a sequence $x_k$ satisfies \refeq{cubic},
then $y_k \equiv (-)^k i x_k$
satisfies $y_{k+1} = f_{-r}(y_k)$.
%
Thus, for an odd $n$, any $n$-cycle
  $\vx{n} = \{x_1, x_2, \dots, x_n, x_1\}$
  of $f_r$
  can be mapped isomorphically to
  an $n$-half-cycle
  $\vy{n} = \{-i x_1, i x_2, \dots, - i x_n, i x_1\}$
  of $f_{-r}$,
  with $\Lambda_r(\vx{n}) = -\Lambda_{-r}(\vy{n})$.
%
So the $n$-half-cycle and $n$-cycle polynomials
  are related by
  $H^*_n(r, \lambda) = P^*_n(-r, -\lambda)$
  for an odd $n$.
%
Similarly, if $n$ is even,
  the $n$-half-cycles and asymmetric $n$-cycles of $f_r$
  can be isomorphically mapped to the corresponding ones of $f_{-r}$,
  with $\Lambda_r(\vx{n}) = \Lambda_{-r}(\vy{n})$.
%
So $H^*_{n}\rlam$,
  $\hat P^*_{n}\rlam$,
  and $A_{n}\rlam$
  are all even polynomials of $r$
\big[but $\deg_r P^*_{n}\rlam$ is even only if $4|n$,
  because if $n/2$ is odd, a symmetric $n$-cycle
  would be mapped to an $(n/2)$-cycle\big].
%
Thus, these polynomials need to be computed only
  at the positive half of $r$ values
  in polynomial interpolation.
%

>>>>

Since the $n$-cycles points constitute
  a subset of the fixed points of $f^n$,
we will first find
  the polynomials of marginal stability
  for the fixed points of $f^n$
  (\refsecs{cyclic}-\ref{sec:examples}).
The polynomials contain
  extraneous factors contributed by short $d$-cycles with $d|n$,
  which are also fixed points of $f^n$.
The factors are removed in \refsec{primfac}.

@@

\refthm{sqrfree}
  shows that any cyclic polynomial can be expanded
  as a combination of the square-free ones.
At the onset and bifurcation points,
  the $\NB$ square-free cyclic polynomials $C_p$ are themselves
  linearly connected by a characteristic equation.

@@

%\begin{theorem}
%  The polynomial $P_n\Rlam$ of all $n$-cycles
%  is a factor of the characteristic polynomial $A_n\Rlam$,
%  and can be computed as
  \begin{equation}
    P_n\Rlam
    = \prod_{cd = n} B_d^c\Rlam^{\mu(c)},
  \label{eq:primfac}
  \end{equation}
where
  $B_d^c\Rlam \equiv \prod_{k=1}^c A_{d}(R, e^{2k\pi i/c} \lambda^{1/c})$
  with
  $\lambda^{1/c}$ being a complex $c$th root of $\lambda$;
Particularly, $B_n^1\rlam = A_n\rlam$.
The outer product runs over all divisors $d$ of $n$.
%\end{theorem}
%
%
%\begin{remark}
%Here, ``minimal'' means that the polynomial contains all zeros.
%It happens to be irreducible for a general $\lambda$.
%At $\lambda = 1$, the polynomial of $R$, however, can be factorized (see later).
%\end{remark}
%
%\begin{remark}
The M\"obius function $\mu(n)$ is $(-1)^k$
  if $n$ is the product of $k$ distinct primes,
  or 0 if $n$ is divisible by a square of a prime;
%Thus,
$\mu(n) = 1, -1, -1, 0, -1, 1, -1, \ldots,$
  from $n = 1$
%The $\mu(n)$ is useful for inversion:
%$g(n) = \sum_{d|n} \mu(n/d) \, h(d)$
%if and only if $h(n) = \sum_{d|n} g(d)$
 \cite{hardy}.
%\end{remark}
%

@@

First, to count the square-free cyclic polynomials,
  we notice a one-to-one mapping
  from the square-free cyclic polynomials
  to the binary necklaces,
%
which are nonequivalent binary $0$-$1$ strings.
%
Two strings are equivalent if they differ only by a circular shift.
%
For example, for $n = 3$, there are $2^3 = 8$ binary strings,
but only four necklaces: 000, 001, 011 and 111,
since 010 and 100 are equivalent to 001,
so are 110 and 101 to 011.
%
%The mapping is constructed as follows.
%
Now given a square-free cyclic polynomial, if its generator contains $x_k$,
  the $k$th character from the right of the necklace is 1,
  otherwise 0,
e.g. the above four necklaces correspond to
  $C_0$, $C_1$, $C_{12}$ and $C_{123}$, respectively.
%For example, if $n = 3$,
% the square-free polynomials for $000$, $100$, $110$ and $111$ are
%  $C_0 = 1$ (generator: 1),
%  $C_1 = x_1 + x_2 + x_3$ (generator: $x_1$),
%  $C_{12} = x_1 x_2 + x_2 x_3 + x_3 x_1$ (generator: $x_1 x_2$)
%  and $C_{123} = x_1 x_2 x_3$ (generator: $x_1 x_2 x_3$), respectively.
Thus,
%
%\begin{theorem}
the number $\NB(n)$ of the square-free cyclic polynomials
  is equal to the number of length-$n$ binary necklaces,
  which is given by \cite{riordan}
\begin{equation}
  %\deg_\lambda A_n\lam =
  \NB(n) = (1/n) \textstyle\sum_{d|n} \varphi(n/d) 2^{\,d},
\label{eq:necklace}
\end{equation}
%
%\label{thm:necklace}
%\end{theorem}
%
where $\varphi(m)$ is Euler's totient function,
  which counts the number of integers from 1 to $m$
  that are coprime to $m$,
  e.g.,
  $\varphi(1) = 1$,
  $\varphi(6) = 2$ for 1 and 5 are coprime to 6 \cite{hardy}.
%
By \refeq{xr},
  $\deg_\lambda A_n\lam = \NB(n)$.
The first few $\NB(n)$ are
2, 3, 4, 6, 8, 14, 20, 36, 60, 108, \dots, % 188, 352, 632, \ldots,
  from $n = 1$.

@@

%\begin{theorem}
Next, we define the $n$-cycle polynomial as
\begin{equation}
  P_n\lam \equiv \textstyle\prod_l [
    \Lambda(\vx{n}^{(l)}) - \lambda
  ],
\notag
\end{equation}
where $\vx{n}^{(l)}$ denotes the $l$th $n$-cycle.
%
Thus, $\deg_\lambda P_n\lam$
  is equal to the number $\LB$ of the $n$-cycles,
  which is given by
  \cite{hao, hao2, lutzky}
\begin{equation}
  %\deg_\lambda P_n\lam =
  \LB(n) = (1/n)\textstyle\sum_{d | n} \mu(n/d) 2^{\,d}.
\label{eq:lyndon}
\end{equation}
%\label{thm:lyndon}
%\end{theorem}
%
%
%\end{proof}
%
The first few $\LB(n)$ are
2, 1, 2, 3, 6, 9, 18, 30, 56, 99, \dots %186, 335, 630, \ldots,
from $n = 1$.
%
%
One can verify that
%Incidentally, $\LB(n)$ is also the number of aperiodic necklaces of length $n$
%(e.g., for $n = 4$,
%  ``$0011$'' is aperiodic, but ``$0101$'' is not, owing to the shorter period 2).
%%
%An $n$-necklace is either aperiodic itself or
%  a repeat of an aperiodic $d$-necklace with $d|n$; so
\begin{equation}
  \NB(n) = \textstyle\sum_{d|n} \LB(d).
\notag
%\label{eq:necklacelyndon}
\end{equation}
%
%Since the left side is the degree in $\lambda$ of $A_n\lam$,
Thus, in order for $A_n\lam$ to accommodate all fixed points of $f^n$,
  each distinct $d$-cycle ($d|n$)
  contributes one and only one zero to $A_n\lam$.
%
%
Comparing the leading coefficient of $A_n\lam$ leads to
%is unity by \refeq{xr}, we must have
%\begin{theorem}
  \begin{equation}
    A_n\lam = \textstyle\prod_{d|n} Q_d^{n/d}\lam,
    \label{eq:prod}
  \end{equation}
where $Q_d^{n/d}(\lambda) \equiv \prod_l
  \big[\Lambda\big(\vx{n}^{(l)}\big) - \lambda\big]$
  with $\vx{n}^{(l)}$ being the $n$ consecutive points
  in the $l$th $d$-cycle,
%
  or the $d$-cycle points $\vx{d}^{(l)}$
  repeated for $c = n/d$ times:
\begin{equation}
  \vx{n}^{(l)}
  =
  \vx{d}^{(l)} \times c
  =
  \{
    x_1^{(l)}, \ldots, x_d^{(l)},
    \; \ldots, \;
    x_1^{(l)}, \ldots, x_d^{(l)}
  \}.
\notag
\end{equation}
%
It follows that
$\Lambda(\vx{n}) = \left[ \Lambda(\vx{d}) \right]^c$.
%
By factoring $\Lambda^c - \lambda$
  in terms of the $c$ complex roots of $\lambda$,
  we get
%
\begin{align}
Q_d^{c}\lam
%  &= \prod_l
%    \Big\{
%      \lambda - \big[\Lambda\big(\vx{d}^{(l)}\big) \big]^{c}
%    \Big\}
%    \notag \\
  &=
    \prod_l
    \left\{
      \prod_{k=1}^c \,
      \left[
          \Lambda\big(\vx{d}^{(l)}\big)
          -
          e^{2k\pi i/c} \lambda^{1/c}
      \right]
    \right\}
    \notag \\
  &=  \prod_{k=1}^c P_d\big(e^{2k\pi i/c} \lambda^{1/c}\big).
\label{eq:QP}
\end{align}
%
%where we have exchanged the order of the two products in the second step.
%
The radical $\lambda^{1/c}$
  does not prevent
  $Q_d^c\lam$ from being a polynomial of $\lambda$,
  thanks to the invariance
  $\lambda \rightarrow e^{2\pi i} \lambda$;
and
  $\deg_\lambda Q_d^c\lam
    = \deg_\lambda P_d\lam = \LB(d)$.

@@

The most time-consuming step of the algorithm in \refsec{algo}
  is the evaluation of the determinant $A_n(R, \lambda = \pm1)$
  in the characteristic equation (\ref{eq:xr}).
%
In implementation,
we found it helpful to use polynomial interpolation
by first evaluating $A_n\Rlam$ at $\deg_R A_n + 1$ values of $R$,
e.g., $R = 0, \pm1, \pm2,\ldots$, then piecing them together
into a polynomial of $R$.
%
Further, using \refeq{primfac}, we only need $\deg_R P_n + 1$ values
  to directly compute $P_n\Rlam$.
%
%
Here,
$\deg_R A_n = n \NB(n)/2$
and $\deg_R P_n = n \LB(n)/2$
  \cite{mira, stephenson1}.
%
%The strategy also allows a trivial parallelization.
%
%\begin{figure}[h]
%  \begin{center}
%  \begin{minipage}{\linewidth}
%        \includegraphics[width=\linewidth]{T14b.png}
%  \end{minipage}%
%  \end{center}
%  \caption{\label{fig:T14b}
%  Coefficients of the 14-cycle bifurcation polynomial in two columns (left to right).
%  The $k$th line from the top is the coefficient before $(4R)^k$ in base-256.
%  The digits are shown by the grayscale from white (0) to black (255).
%  The leading bar before each line is the sign (black: $-$, white: $+$).
%  }
%\end{figure}

We coded the algorithm into a Mathematica program,
  and computed the polynomials for $n$ up to 14.
The data are saved on the website http://logperiod.appspot.com.
%
%The bifurcation polynomial of the 14-cycle in shown in \reffig{T14b}.
%
%The exact polynomials of these maps are generally too large to print on paper,
%  e.g., the bifurcation polynomial of the 14-cycle
%    takes about 27 megabytes.
%
%The representative $r$ values are listed in \reftab{rval}.
%
%Compared with other approaches \cite{hitzl, gordon, bailey2, kk1, hao, hao2}
%it appeared to be simpler and more transparent,
%as it only requires the computation of a determinant.
%
As shown in \reffig{logtime}, the algorithm
  outperformed the resultant \cite{burm} and
  Gr\"obner basis \cite{kk1} methods,
  by at least two orders of magnitudes.
%
This is unsurprising,
  because the resultant of $f^n(x)-x$ and $\frac{d}{dx}f^n(x)-\lambda$
  is the determinant of the Sylvester matrix, which has
  $2^{n+1}-1$ rows and columns
  [as $\deg_x f^n(x) = 2^n$],
  while the matrix in \refeq{xr}
  is of only $\NB \approx 2^n/n$ [cf. \refeq{necklace}] rows and columns.
%Thus, the current algorithm is expected to be faster.

@@

Finally, we can further reduce the computational cost
  by half for an even $n$.
%
If a sequence $x_k$ satisfies \refeq{cubic},
then $y_k \equiv (-)^k i x_k$
satisfies $y_{k+1} = f_{-r}(y_k)$.
%
Thus, for an odd $n$, any $n$-cycle
  $\vx{n} = \{x_1, x_2, \dots, x_n, x_1\}$
  of $f_r$
  can be mapped isomorphically to
  an $n$-half-cycle
  $\vy{n} = \{-i x_1, i x_2, \dots, - i x_n, i x_1\}$
  of $f_{-r}$,
  with $\Lambda_r(\vx{n}) = -\Lambda_{-r}(\vy{n})$.
%
So the $n$-half-cycle and $n$-cycle polynomials
  are related by
  $H^*_n(r, \lambda) = P^*_n(-r, -\lambda)$
  for an odd $n$.
%
Similarly, if $n$ is even,
  the $n$-half-cycles and asymmetric $n$-cycles of $f_r$
  can be isomorphically mapped to the corresponding ones of $f_{-r}$,
  with $\Lambda_r(\vx{n}) = \Lambda_{-r}(\vy{n})$.
%
So $H^*_{n}\rlam$,
  $\hat P^*_{n}\rlam$,
  and $A_{n}\rlam$
  are all even polynomials of $r$
\big[but $\deg_r P^*_{n}\rlam$ is even only if $4|n$,
  because if $n/2$ is odd, a symmetric $n$-cycle
  would be mapped to an $(n/2)$-cycle\big].
%
Thus, these polynomials need to be computed only
  at the positive half of the $r$ values
  in polynomial interpolation.
%
